---
title: "GER-GSR"
author: "Franklin"
date: "6/25/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Loading packages 

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
library(tidyverse)
library(here)
library(readxl)
library(viridis)
library(ggthemes)
library(lme4)
library(cowplot)
library(broom)
library(agricolae)
library(lattice)
library(influence.ME)
library(emmeans)
library(scales)
```


## Gibberella Stalk Rot trials - GSR

### Reading data

```{r}
dat_stalk_2017 <- read_excel(here("stalk-data", "data", "data_stalk.xlsx"), 1) %>% 
  subset(year == "2017") %>% 
  mutate(isolate = as.factor(isolate)) %>% 
  group_by(year, hybrid, host, species, isolate,  block) %>%
  summarise(sev_plant = mean(sev, rm.na = T)) %>% 
  select(year, hybrid, host, species, isolate, stalk_half = block, sev_plant) 



dat_stalk_2018 <- read_excel(here("stalk-data", "data", "data_stalk.xlsx"), 1) %>% 
  subset(year == "2018") %>% 
  mutate(isolate = as.factor(isolate)) %>% 
  group_by(year, hybrid, host, species, isolate,  stalk_half) %>%
  summarise(sev_plant = mean(sev, rm.na = T)) 




dat_stalk <- bind_rows(dat_stalk_2017, dat_stalk_2018) %>% 
  filter(isolate != "test" &
           sev_plant != "NA")%>% 
  ungroup() %>% 
  mutate(isolate = dplyr::recode(isolate, 
  "1" = "188",
  "2" = "189",
  "3" = "190",
  "5" = "191",
  "6" = "192",
  "7" = "193",
  "9" = "194",
  "11" = "195",
  "12" = "196",
  "14" = "197",
  "16" = "1",
  "17" = "2",
  "18" = "3",
  "19" = "4",
  "22" = "5",
  "26" = "198",
  "27" = "6",
  "28" = "7",
  "29" = "199",
  "30" = "8",
  "31" = "9",
  "33" = "10",
  "35" = "11",
  "36" = "12",
  "37" = "13",
  "39" = "14",
  "40" = "15",
  "43" = "16",
  "44" = "17",
  "45" = "18",
  "47" = "200",
  "48" = "201",
  "49" = "202",
  "50" = "203",
  "51" = "204",
  "55" = "205",
  "56" = "206",
  "58" = "207",
  "59" = "208",
  "60" = "209",
  "21" = "21")) %>% 
  mutate(host1 = paste(species, host, sep = "_")) 

```

## Plots
### Stalk by hybrid

Plotting GSR severity by species and hybrids. 


```{r}
dat_stalk %>% 
  filter(species != "test") %>% 
  group_by(isolate, species, host, year,hybrid) %>% 
  summarise(sev_plant = mean(sev_plant)) %>% 
  ggplot(aes(species, sev_plant))+
  geom_boxplot(aes(fill = host), size = 0.6, 
               outlier.colour = NA, width=0.35, 
               position = position_dodge(width = 0.5)) + 
  scale_fill_grey(name = "", start = 1, end = 1)+
  geom_jitter(position=position_jitterdodge(dodge.width=0.7), alpha=0.6, aes(colour=host))+
  #coord_flip()+
  facet_grid(~year*hybrid) +
  #scale_colour_discrete(breaks=c("corn", "wheat"),
  #  labels = c("milho", "trigo"))+
  theme_minimal()+ 
  #ylim(0,80)+
  theme(legend.position = "top",
        panel.grid.major.x = element_blank(), 
        plot.margin = unit(c(-0.6, 0.1, -0.2, 0.1), "cm"),
        legend.margin=margin(0,0,0,0),
        legend.box.margin=margin(-0,-10,-10,-10))+
 labs(y = "GSR severity (%)" ,
      x = "", title = "",
      color = "")+
  ggsave("figs/fig2_stalk.png", width=6, height=2.8, dpi=300)


```
```{r}
dat_stalk %>% 
  filter(species != "test") %>% 
  summary(.$sev_plant)
```


## Mixed model estimates

   We will use the `lmer` function of the `lme4` to fit the mixed models. We log-transformed `sev`data (log[sev+1]) so variance could be reduced.

Let’s fit a mixed model for the GER data.

```{r, message=FALSE, warning=FALSE}
mix1 <- lmer (log(sev_plant+1) ~ species * host + (1|hybrid/host/isolate), data=dat_stalk, REML=FALSE)

plot(mix1, type = c("p", "smooth"), col = "black")
qqmath(mix1, id = 0.05, col = "black")

```


```{r}
library(car)
Anova(mix1)


```
There is no significant interaction. Thus, we just test the effect of species. 

```{r}

mix2 <- lmer (log(sev_plant+1) ~ species + (1|hybrid/host/isolate), data=dat_stalk, REML=FALSE)

plot(mix2, type = c("p", "smooth"), col = "black")
qqmath(mix2, id = 0.05, col = "black")

```

```{r}
Anova(mix2)

medias_species <- emmeans(mix2, pairwise ~ species, type = "response")
medias_species

```

### LSmeans

```{r}
gsr_ic <- emmeans(mix2,  ~ species, type = "response") %>% as.data.frame() %>%
  ggplot(aes(species, response, colour = species))+
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = lower.CL, ymax= upper.CL, width = 0))+
  #facet_grid(~host)+
  theme_minimal()+
  ylim(0,70)+ 
  theme(legend.position="none",
        panel.grid.major.x = element_blank() 
        #axis.text.x = element_text(angle = 45, hjust = 1),
        #legend.text = element_text(face = "italic", size = 6),
        #strip.text = element_text(face = "italic"),
        #plot.margin = margin(0, 0.1, 0.1, 0.1, unit = "cm")
       )+
 labs(y = "GSR severity (%)", x ="" )#+
  #ggsave("figs/ic-emmeans_GSR.png", width=2.3, height=3)

gsr_se  <- emmeans(mix2,  ~ species, type = "response") %>% as.data.frame() %>%
  ggplot(aes(species, response, colour = species))+
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = response-SE, ymax= response+SE, width = 0))+
  #facet_grid(~host)+
  theme_minimal()+
  ylim(-1,30)+ 
  theme(legend.position="none",
        panel.grid.major.x = element_blank(), 
        #axis.text.x = element_text(angle = 45, hjust = 1),
        #legend.text = element_text(face = "italic", size = 6),
        #strip.text = element_text(face = "italic"),
        plot.margin = margin(0.85, 0.1, 0.17, 0.1, unit = "cm")
       )+
 labs(y = "GSR severity (%)", x ="" )#+
  #ggsave("figs/SE-emmeans_GSR.png", width=2.3, height=3)
  
```


## Gibberella Ear Rot trials - GER

### Reading data

```{r}
dat_2017 <- read_excel(here("ear-data", "data_ear.xlsx"), 1) %>% 
  mutate(year = "2017") %>% 
  gather(hybrid, sev, 6:7) %>% 
  filter(sev != "NA") %>% 
  mutate(sev = as.numeric(sev)) %>% 
  group_by(year, hybrid, block, species, host, isolate) %>% 
  summarise(sev = mean(sev, na.rm = TRUE)) %>% 
  select(year, hybrid, species, host, isolate, ear = block, sev) 
  
  
  
dat_2018 <- read_excel(here("ear-data", "data_ear.xlsx"), 2) %>% 
  mutate(year = "2018") %>% 
  gather(hybrid, sev, 3:4) %>% 
  filter(sev != "NA") %>% 
  select(year, hybrid, species, host, isolate, ear, sev) 
  #group_by(year, hybrid, species, host, isolate) %>% 
  #summarise(sev = mean(sev, na.rm = TRUE)) %>% 
  #select(year, hybrid, species, host, isolate, sev) %>% 
  #ungroup()
 


dat_ear <- bind_rows(dat_2017, dat_2018) %>% 
  filter(isolate != "test" & sev != 0 ## remove the ears with zero severity 
         )  %>% 
  mutate(isolate = dplyr::recode(isolate, 
  "1" = "188",
  "2" = "189",
  "3" = "190",
  "5" = "191",
  "6" = "192",
  "7" = "193",
  "9" = "194",
  "11" = "195",
  "12" = "196",
  "14" = "197",
  "16" = "1",
  "17" = "2",
  "18" = "3",
  "19" = "4",
  "22" = "5",
  "26" = "198",
  "27" = "6",
  "28" = "7",
  "29" = "199",
  "30" = "8",
  "31" = "9",
  "33" = "10",
  "35" = "11",
  "36" = "12",
  "37" = "13",
  "39" = "14",
  "40" = "15",
  "43" = "16",
  "44" = "17",
  "45" = "18",
  "47" = "200",
  "48" = "201",
  "49" = "202",
  "50" = "203",
  "51" = "204",
  "55" = "205",
  "56" = "206",
  "58" = "207",
  "59" = "208",
  "60" = "209",
  "21" = "21")) %>% 
  ungroup() %>% 
  mutate(ear = as.factor(ear),
         isolate = as.factor(isolate)) %>% 
  mutate(host1 = paste(species, host, sep = "_"))

summary(dat_ear$sev)


```

## Plots


```{r}
dat_ear %>% 
  #filter(hybrid == "supremo") %>% 
  #group_by(isolate, specie, host, block) %>% summarise(mean_sev = mean(sev_plant)) %>% 
  ggplot(aes(hybrid, sev))+
  geom_boxplot(size = 0.2, outlier.shape = NA) + 
  geom_jitter(width=0.1, size=1.5,  alpha=0.3, aes(color=species))+
  #coord_flip()+
  facet_wrap(~isolate) +
  #scale_colour_discrete(breaks=c("corn", "wheat"),
  #  labels = c("milho", "trigo"))+
  theme_minimal()+ 
  #ylim(0,80)+
  theme(legend.position = "top",axis.text.x = element_text(angle = 45, hjust = 1)
        #,plot.title = element_text(face="bold")
        )+
 labs(y = "Severity (%)" ,
      x = "", title = "",
      color = "")
```

### Ear - by species

```{r}
dat_ear %>% 
  filter(species != "test") %>% 
  group_by(isolate, species, host, year,hybrid) %>% 
  summarise(sev = mean(sev)) %>% 
  ggplot(aes(species, sev))+
  geom_boxplot(aes(fill = host), size = 0.6, 
               outlier.colour = NA, width=0.35, 
               position = position_dodge(width = 0.5)) + 
  scale_fill_grey(name = "", start = 1, end = 1)+
  geom_jitter(position=position_jitterdodge(dodge.width=0.7),  alpha=0.6, aes(color=host))+
  #coord_flip()+
  facet_wrap(~year*hybrid,  ncol = 4) +
  #scale_colour_discrete(breaks=c("corn", "wheat"),
  #  labels = c("milho", "trigo"))+
  theme_minimal()+ 
  #ylim(0,80)+
  theme(legend.position = "top",
        panel.grid.major.x = element_blank(), 
        plot.margin = unit(c(-0.6, 0.1, -0.2, 0.1), "cm"),
        legend.margin=margin(0,0,0,0),
        legend.box.margin=margin(-0,-10,-10,-10))+
 labs(y = "GER severity (%)" ,
      x = "", title = "",
      color = "")+
  ggsave("figs/fig3_ear_sem_zeros.png", width=6, height=2.8, dpi=300)




```

```{r}
dat_ear %>% 
  filter(species != "test") %>% 
  summary(.)
```

## Mixed model estimates

  We will use the `lmer` function of the `lme4` to fit the mixed models. We log-transformed `sev`data (log[sev+1]) so variance could be reduced.

Let’s fit a mixed model for the GER data. 

```{r, message=FALSE, warning=FALSE}
mix0 <- lmer (log(sev+1) ~ species * host + (1|hybrid/host/isolate), data=dat_ear, REML=FALSE)
mix1 <- lmer (log(sev+1) ~ species * host + (1|hybrid/isolate), data=dat_ear, REML=FALSE)

anova(mix0, mix1) ## Going to use mix1 model. Keeping host random effect give me problems of convergence once its fixed effect is already explaining everything. 


plot(mix1, type = c("p", "smooth"), col = "black")
qqmath(mix1, id = 0.05, col = "black")

```


Once we have our best fit model, let's first test the effect of the interaction.

```{r}
library(car)
Anova(mix1)


```
There is significant interaction. THus, we create different data sets for each hybrid in order to test the effect of species within each host of origin. 





```{r}

medias_species <- emmeans(mix1,  ~ species*host, type = "response" )
summary(medias_species, by = "species")
summary(medias_species, by = "host")


```
```{r}
# checking significant interaction within each factor.
pairs(medias_species, by = "species", type = "response" )

pairs(medias_species, by = "host", type = "response" )


```


## LSmeans

```{r}
ger_ic <- medias_species %>% as.data.frame() %>%
  ggplot(aes(species, response, colour = species))+
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = lower.CL, ymax= upper.CL, width = 0))+
  facet_grid(~host)+
  theme_minimal()+
  ylim(-1,90)+ 
  theme(legend.position="none",
        panel.grid.major.x = element_blank() 
        #axis.text.x = element_text(angle = 45, hjust = 1),
        #legend.text = element_text(face = "italic", size = 6),
        #strip.text = element_text(face = "italic"),
        #plot.margin = margin(0, 0.1, 0.1, 0.1, unit = "cm")
       )+
 labs(y = "GER severity (%)", x ="" )#+
  #ggsave("figs/ic-emmeans_GER_semzeros.png", width=3, height=3)

ger_se <- medias_species %>% as.data.frame() %>%
  ggplot(aes(species, response, colour = species))+
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = response-SE, ymax= response+SE, width = 0))+
  facet_grid(~host)+
  theme_minimal()+
  ylim(-1,30)+ 
  theme(legend.position="none",
        panel.grid.major.x = element_blank() 
        #axis.text.x = element_text(angle = 45, hjust = 1),
        #legend.text = element_text(face = "italic", size = 6),
        #strip.text = element_text(face = "italic"),
       #plot.margin = margin(0, 0, 0, 0, unit = "cm")
       )+
 labs(y = "GER severity (%)", x ="" )#+
  #ggsave("figs/SE-emmeans_GER_semzeros.png", width=3, height=3)
  
```


## Combo plot


```{r}
plot_grid(ger_se, gsr_se, labels=c('A', 'B'), align = "hv", ncol=2)+
ggsave("figs/fig4_combo_SE_semzeors.png",  width=5.8, height=3, dpi=300)

plot_grid(ger_ic, gsr_ic, labels=c('A', 'B'), align = "hv", ncol=2)+
ggsave("figs/fig4_combo_ic_semzeors.png",  width=5.8, height=3, dpi=300)

```



#Correlation 
```{r}

ear <- dat_ear %>% 
  filter(species != "test") %>% 
  group_by(isolate, species, host, hybrid) %>% 
  summarise(sev_ear = mean(sev))


stalk <- dat_stalk %>% 
  filter(species != "test") %>% 
  group_by(isolate, species, host, hybrid) %>% 
  summarise(sev_stalk = mean(sev_plant)) 



stalk_ear <- left_join(stalk, ear)




M <- cor.test(stalk_ear$sev_stalk, stalk_ear$sev_ear, use = "pairwise.complete.obs", method = "pearson")

M 


stalk_ear %>% 
  #filter(hybrid == "supremo") %>% 
  ggplot(aes(sev_ear, sev_stalk))+
  geom_point( size = 3, alpha=0.7
              , aes(color = species)
              )+
  geom_smooth(method="lm", fullrange = TRUE, se = F, size = 0.7, color = "black")+
  #geom_text(x = 25, y = 90, label = lm_eqn(stalk_ear), parse = TRUE, size = 2)+
  facet_grid (~ hybrid, scales = "free") +
  #scale_fill_grey(start = 0.8, end = 0.5)+
  theme_minimal()+ 
  ylim(0,100)+xlim(0,100)+
  theme(legend.position = "top"
        #,plot.title = element_text(face="bold")
        )+
  labs(x = "Severidade em colmos (%)",
      y = "Severidade em colmos (%)", 
      #title = "SUPREMO",
      color = "")+  
  ggsave("stalk_ear_correlation.png", width=4, height=3.5)


lm_eqn <- function(stalk_ear){
    fit <- lm(sev_stalk ~ sev_ear, stalk_ear);
    eq <- substitute(italic(y) == a + b %.% italic(x)*","~~italic(r)^2~"="~r2*","~~italic(P)~"<"~ p, 
         list(a = format(coef(fit)[1], digits = 2), 
              b = format(coef(fit)[2], digits = 2), 
             r2 = format(summary(fit)$r.squared, digits = 3),
             p = "0.001"))
    as.character(as.expression(eq));                 
}

summary(fit)

```

# Mycotoxin 


```{r}

dat_maize1 <- read_excel(here("ear-data", "FY19_Vaillancourt_Complete_Results_11_26_19.xlsx"), 3) %>%
  select(-"Lab_ID#") %>% 
  mutate(exp = "1")

dat_maize2 <- read_excel(here("ear-data", "FY19_Vaillancourt_Complete_Results_11_26_19.xlsx"), 4) %>%
  select(-"Lab_ID#") %>% 
  mutate(exp = "2")

dat_mycot_maize <- dat_maize2 %>% 
  #bind_rows(., dat_maize2) %>%
  filter(strain != "test" &
           strain != "NA" ) %>% 
  arrange(strain) %>% 
  gather(mycotoxin, ppm, 5:9) #%>% 
  #filter(mycotoxin != "ZON")

#openxlsx::write.xlsx(dat_mycot_wheat, "dat_mycot_wheat.xlsx", colNames = TRUE)




```


```{r}
myColors1 <- c("#00AFBB","blue",  "#E7B800", "#FC4E07", "#2CA25F")
names(myColors1) <- levels(factor(dat_mycot_maize$mycotoxin, levels=c("15ADON", "3ADON", "DON", "NIV", "ZON")))
colScale1 <- scale_fill_manual(name = "", values = myColors1)

fgra_maize <-dat_mycot_maize %>% 
  filter(species == "Fgra" ) %>% 
  group_by(mycotoxin, host, species, strain) %>% 
  summarize(ppm = mean(ppm, na.rm = T)) %>% 
  ggplot()+
  geom_bar(aes(y=ppm, x=factor(strain),  fill = mycotoxin), stat = "identity", alpha =0.8, width = 0.6)+
 theme_minimal()+
 facet_wrap (species ~ host, scales = "free_x") +
 colScale1+
  #scale_fill_manual(labels = c("15ADON", "3ADON", "DON", "NIV", "ZON"), values=c("#00AFBB","blue",  "#E7B800", "#FC4E07", "#2CA25F"))+ #"#00AFBB", "#E7B800", "#FC4E07"
  theme(legend.position = "none",
        legend.margin=margin(0,0,0,0),
        legend.box.margin=margin(-0,-10,-10,-10),
        plot.margin = unit(c(0.1, 0.1, -0.2, 0.1), "cm"),
        axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
        panel.grid.major.x = element_blank() 
        )+
 scale_y_continuous(expand=c(0,0))+
 labs(y =  (expression(paste('Trichothecenes', ' (', mu,'g/g)'))), fill ="", x = "")#+
  #ggsave("figs/tric_rice.png", width=4.2, height=2.8, dpi=300)


fmer_maize <- dat_mycot_maize %>% 
  filter(species == "Fmer" ) %>% 
  group_by(mycotoxin, host, species, strain) %>% 
  summarize(ppm = mean(ppm, na.rm = T)) %>% 
  ungroup() %>% 
  ggplot()+
  geom_bar(aes(y=ppm, x=factor(strain),  fill = mycotoxin), stat = "identity", alpha =0.8, width = 0.6)+
 theme_minimal()+
 facet_wrap (species ~ host, scales = "free_x") +
 colScale1+
  theme(legend.position = "none",
        legend.margin=margin(0,0,0,0),
        legend.box.margin=margin(-0,-10,-10,-10),
        plot.margin = unit(c(0.1, 0.1, -0.2, 0.1), "cm"),
        axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
        panel.grid.major.x = element_blank() 
        )+
 scale_y_continuous(expand=c(0,0))+
 labs(y =  "", fill ="", x = "")#+
  #ggsave("figs/tric_rice.png", width=4.2, height=2.8, dpi=300)

```

## Plots grids

```{r}
legend_maize <- get_legend(dat_mycot_maize %>% ggplot()+
  geom_bar(aes(y=ppm, x=reorder(strain,ppm),  fill = mycotoxin), stat = "identity", alpha =0.8, width = 0.6)+
 #theme_minimal()+
 colScale1 + theme(legend.position="top")+guides(color = guide_legend(nrow = 1)))

grid1 <- plot_grid(fgra_maize, fmer_maize, labels=c('A', ''), align = "hv", ncol=2)
grid3<-plot_grid(legend_maize, grid1, align = "hv", ncol = 1, rel_heights = c(.1, 1))+
ggsave("figs/fig6_mycot_maize2.png",  width=10.5, height=3.3, dpi=300)
grid3
```



### Descritive analysis

```{r}
datx <- dat_mycot_maize %>% 
  filter(species == "Fmer" ) %>% 
  group_by(mycotoxin, host, species, strain) %>% 
  summarise(ppm = mean(ppm, na.rm = T)) %>% 
  filter(ppm != "NaN")
summary(datx)
table(datx$host, datx$species, datx$mycotoxin)

#dat1 <- dat_rice %>% 
 # filter(species != "Fgra_test") %>% 
  #filter(mycotoxin == "ZON")
  
  
#cohen.d(dat1$ppm ~ dat1$species, na.rm=TRUE)

dat_mycot_maize %>% 
  #filter(host == "maize") %>% 
  group_by(mycotoxin, host, species) %>% 
  summarise(mean = round(mean(ppm, na.rm =TRUE), 2), 
            n = length(ppm),
            se = round(sd(ppm, na.rm =TRUE)/sqrt(n), 2),
            sd = round(sd(ppm, na.rm =TRUE),2))%>% 
  filter(mean != "NaN") %>% 
  as.data.frame()
```




## sev x micotoxina 

```{r, warning=FALSE}

fmer_820 <- dat_ear %>% 
  filter(hybrid == "BM820") %>% 
  filter(species == "Fmer" ) %>% 
  group_by(hybrid,  species, host,  isolate) %>% 
  summarise(mean_sev = mean(sev, na.rm = T),
            sd_sev = sd(sev, na.rm=TRUE),
           n = length(sev),
           se_sev = sd_sev / sqrt(n)) %>% 
  ggplot(aes(isolate, mean_sev))+
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = mean_sev-se_sev, ymax= mean_sev+se_sev, width = 0))+
  theme_minimal()+
  facet_wrap (species ~ host, scales = "free_x") +
 theme(legend.position = "none",
        legend.margin=margin(0,0,0,0),
        legend.box.margin=margin(-0,-10,-10,-10),
        plot.margin = unit(c(0.1, 0.1, -0.2, 0.1), "cm"),
        axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
        panel.grid.major.x = element_blank() 
        )+
 ylim(c(0,100))+
 #scale_y_continuous(expand=c(0,0))+
 labs(y = "", x ="" )


fgra_820 <- dat_ear %>% 
  filter(hybrid == "BM820") %>% 
  filter(species == "Fgra" ) %>% 
  group_by(hybrid,  species, host,  isolate) %>% 
  summarise(mean_sev = mean(sev, na.rm = T),
            sd_sev = sd(sev, na.rm=TRUE),
           n = length(sev),
           se_sev = sd_sev / sqrt(n)) %>% 
  ggplot(aes(isolate, mean_sev))+
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = mean_sev-se_sev, ymax= mean_sev+se_sev, width = 0))+
  theme_minimal()+
  facet_wrap (species ~ host, scales = "free_x") +
 theme(legend.position = "none",
        legend.margin=margin(0,0,0,0),
        legend.box.margin=margin(-0,-10,-10,-10),
        plot.margin = unit(c(0.1, 0.1, -0.2, 0.1), "cm"),
        axis.text.x = element_text(angle = 90, hjust = 0, vjust = 0.5),
        panel.grid.major.x = element_blank() 
        )+
  ylim(c(0,100))+
 #scale_y_continuous(expand=c(0,0))+
 labs(y = "GER severity (%)", x ="" )

```

## grid
```{r}

grid2 <- cowplot::plot_grid(fgra_820, fmer_820, labels=c('B', ''), align = "hv", ncol=2)

cowplot::plot_grid(grid3, grid2, labels=c('', ''), align = "hv", ncol=1)+
ggsave("figs/fig_combo1.png",  width=7.7, height=4.3, dpi=300)

```
### Corr NIV
```{r}

fmer_820 <- dat_ear %>% 
  filter(hybrid == "BM820") %>% 
  filter(species == "Fmer" ) %>% 
  group_by(hybrid,  species, host,  isolate) %>% 
  summarise(mean_sev = mean(sev, na.rm = T))
  
  
fmer_maize <- dat_mycot_maize %>% 
  filter(species == "Fmer" &
         strain != "NA") %>% 
  group_by(mycotoxin, host, species, isolate=strain) %>% 
  summarize(ppm = mean(ppm, na.rm = T)) %>% 
  filter(mycotoxin == "NIV" ) 
  
  

corr <- left_join(fmer_820, fmer_maize) 
  #filter() %>% 
  #select(-c(isolate, species, host, hybrid, mycotoxin)) 



niv_cor<- cor.test(corr$mean_sev, corr$ppm, use = "pairwise.complete.obs", method = "pearson")
niv_cor 
```


### Corr NIV
```{r}

fgra_820 <- dat_ear %>% 
  filter(hybrid == "BM820") %>% 
  filter(species == "Fgra" ) %>% 
  group_by(hybrid,  species, host,  isolate) %>% 
  summarise(mean_sev = mean(sev, na.rm = T))
  
  
fgra_maize <- dat_mycot_maize %>% 
  filter(species == "Fgra" &
         strain != "NA") %>% 
  group_by(mycotoxin, host, species, isolate=strain) %>% 
  summarize(ppm = mean(ppm, na.rm = T)) %>% 
  filter(mycotoxin == "DON" ) 
  
  

corr <- left_join(fgra_820, fgra_maize) 
  #filter() %>% 
  #select(-c(isolate, species, host, hybrid, mycotoxin)) 



don_cor<- cor.test(corr$mean_sev, corr$ppm, use = "pairwise.complete.obs", method = "pearson")
don_cor 
```

## Multivariate

### Preparing datasets

```{r warning=FALSE, include=FALSE}
stalk_multivar <- dat_stalk %>%
  mutate(var = "stalk") %>% 
  unite(hyb, var, hybrid, sep = "_") %>% 
  group_by(hyb, host, species, strain = isolate) %>%
  summarize(sev_stalk = mean(sev_plant, na.rm = TRUE)) %>% 
  spread(hyb, sev_stalk) 

ear_multivar <- dat_ear %>%
  mutate(var = "ear") %>% 
  unite(hyb, var, hybrid, sep = "_") %>% 
  group_by(hyb, host, species, strain = isolate) %>%
  summarize(sev_ear = mean(sev, na.rm = TRUE)) %>% 
  spread(hyb, sev_ear)

corn_aggress <- ear_multivar %>% 
  left_join(., stalk_multivar) %>% 
  #filter(strain != "Fm14" &
   #      strain != "Fm15" &
    #     strain != "Fm16" &
     #    strain != "Fm21"&
      #   strain != "Fg190" &
       #  strain != "Fg196" &
        # strain != "Fg196") %>% 
  arrange(strain)


dat_multivar <- read_excel( "dat_multivar.xlsx", 1) %>% 
  left_join(., stalk_multivar) %>% 
  left_join(., ear_multivar)



table(corn_aggress$species, corn_aggress$host)
#openxlsx::write.xlsx(corn_aggress, "aggress_corn.xlsx", colNames = TRUE)
```

# FACTOR mine package

```{r}
#http://www.sthda.com/english/articles/31-principal-component-methods-in-r-practical-guide/112-pca-principal-component-analysis-essentials/
library("FactoMineR")
mult1 <- corn_aggress

mult <- corn_aggress %>% 
  select(-c(species, host)) 
  

rownames(mult) <- mult$strain ## set rownames
mult <- mult %>% select(-strain)  ## remove first column



res.pca <- PCA(mult, graph = FALSE)

print(res.pca)


```

```{r}
library("factoextra")
eig.val <- get_eigenvalue(res.pca)
eig.val

fviz_eig(res.pca, addlabels = TRUE, ylim = c(0, 50))
```

## PCA plot

```{r}
fviz_pca_biplot(res.pca, 
                # Individuals
                geom.ind = "point",
                #geom.var = c("point", "text"),
                fill.ind = mult1$species, #col.ind = "black",
                pointshape = 21, 
                pointsize = "contrib",
                mean.point = FALSE, #remove group mean point
                #palette = "jco",
                palette = c("#00AFBB", "#E7B800", "#FC4E07"),
                addEllipses = TRUE, #ellipse.type = "confidence",
                # Variables
                alpha.var ="contrib", 
                col.var = "contrib",
                gradient.cols = "RdYlBu",
                invisible = "var",
                legend.title = list(fill = "Species", color = "Contrib",
                                    alpha = "Contrib", size = "Contrib")
                )+
  labs(title = "")+ 
  xlim(c(-5.2, 5.23))+
  ylim(c(-5.2, 5.2))+ #scale_fill_brewer(palette="Dark2") +
  ggsave("figs/fig6_biplot.png", width=6, height=4.8, dpi=300)
  

fviz_pca_biplot(res.pca, 
                # Individuals
                geom.ind = "point",
                #geom.var = c("point", "text"),
                fill.ind = mult1$species, #col.ind = "black",
                pointshape = 21, 
                pointsize = "contrib",
                mean.point = FALSE, #remove group mean point
                #palette = "jco",
                #palette = c("#00AFBB", "#E7B800", "#FC4E07"),
                addEllipses = TRUE, #ellipse.type = "confidence",
                # Variables
                #alpha.var ="contrib", 
                #col.var = "contrib",
                gradient.cols = "RdYlBu",
                invisible = "var",
                legend.title = list(fill = "Species", color = "Species", size = "Contrib")
                )+
  labs(title = "")+ 
  xlim(c(-5.2, 5.25))+
  ylim(c(-5.2, 5.2))+ #scale_fill_brewer(palette="Dark2") +
  ggsave("figs/biplot1.png", width=6, height=4.8, dpi=300)

```

### Componentes principais

```{r}

#attach (dat4)

# Matriz R
r = cor (dcomp [, 4:11])
r

cpr = princomp (dcomp [, 4:11], cor = T)
summary(cpr, loading = T)


# Extraindo os autovetores dos componentes principais 1 à 4
autovetores=cpr$loadings[,1:8]
autovetores


# Extraindo os desvios-padrão dos componentes principais = Autovalores

sdev=data.frame((cpr$sdev)^2)
sdev


# Grau de explicação dos k CP
grau1=sdev[1,1]/sum(sdev)*100
grau1
grau2=sdev[2,1]/sum(sdev)*100
grau2
grau3=sdev[3,1]/sum(sdev)*100
grau3


grau_expli=sum(grau1, grau2,grau3)
grau_expli 


# Obtendo os scores dos componentes principais
cp1r = cpr$scores [, 1]
cp2r = cpr$scores [, 2]


cp_data <- bind_cols(dcomp,cp1r)
plot (cp1r, cp2r)
text (cp1r, cp2r, dcomp$isolate)




rycp1r = cor (dcomp [, 4:11], cp1r)
rycp2r = cor (dcomp [, 4:11], cp2r)


# Data.frame com os scores gerais dos componentes principais
Cor_YCP<-data.frame(rycp1r,rycp2r)
Cor_YCP



det(r)
x2=-(87-1-((2*10+5)/6))*(log(det(r)))
x2

qchisq(0.95, df=45) # ((10^2-10)/2)
head(dat4)
```

### CP plot



```{r}
library(ggrepel)

dcomp2 <- cbind(dcomp, cp1r, cp2r)

dcomp2 %>% 
  #filter(hybrid == "supremo") %>% 
  #group_by(species, strain, rep) %>% 
  #summarise(mean_cm = mean(cm, na.rm = T)) %>% 
  ggplot(aes(cp1r, cp2r))+
  #geom_point()
  #geom_boxplot(size = 0.2, outlier.shape = NA) + 
  geom_point( size = 3, alpha=0.7
              , aes(shape = perit_group, color = species)
              )+
  geom_vline(xintercept = 0)+
  geom_hline(yintercept = 0)+
  #coord_flip()+
 # scale_colour_discrete(breaks=c("maize", "wheat"),
  #  labels = c("milho", "trigo"))+
  #facet_grid (~ specie, scales = "free") +
  #scale_fill_grey(start = 0.8, end = 0.5)+
  theme_minimal()+ 
  ylim(-4,4)+xlim(-4,4)+
  theme(legend.position = "top"
        #,plot.title = element_text(face="bold")
        )+
  geom_label_repel(aes(label = strain, size=2),
                  box.padding   = 0.35, 
                  point.padding = 0.5,
                  segment.color = 'grey50')+
 labs(y = "First PC 22.93% of variance",
      x = "Second PC 20.64% of variance", 
      #title = "SUPREMO",
      color = "", shape = "Perit.")+  
  ggsave("figs/princ_components.png", width=5, height=5.5)


```

###cluster analysis


```{r}
# obtem os valores medios

z1 = (dcomp$spore_conc_105 - mean (dcomp$spore_conc_105)) / sd (dcomp$spore_conc_105)
z2 = (dcomp$mean_cm - mean (dcomp$mean_cm)) / sd (dcomp$mean_cm)
z3 = (dcomp$stalk_kws - mean (dcomp$stalk_kws)) / sd (dcomp$stalk_kws)
z4 = (dcomp$stalk_sup  - mean (dcomp$stalk_sup )) / sd (dcomp$stalk_sup)
z5 = (dcomp$ear_kws - mean (dcomp$ear_kws)) / sd (dcomp$ear_kws)
z6 = (dcomp$ear_sup - mean (dcomp$ear_sup)) / sd (dcomp$ear_sup)
z7 = (dcomp$ear_MG580PW - mean (dcomp$ear_MG580PW)) / sd (dcomp$ear_MG580PW)
z8 = (dcomp$ear_BM820 - mean (dcomp$ear_BM820)) / sd (dcomp$ear_BM820)

dageap = cbind (dcomp, z1, z2, z3, z4, z5, z6, z7, z8)



dep = dist (dageap [, 14:21], method = "euclidean")
agr1 = hclust (dep, method = "single")


plot (agr1, hang = -1, labels = dcomp$strain, xlab = "", ylab = "Distance")
agr1$height
mcof = cophenetic (agr1)
mcof
cor (dep, mcof)
mojena = mean (agr1$height) + 1.25 * sd (agr1$height)
mojena
k = length (agr1$height [agr1$height > mojena]) + 1
abline (h = mojena)
k
rect.hclust (agr1, k = 5)
grupo = cutree (agr1, k = 5)
dageapg = cbind (dcomp, grupo)
dageapg
#tapply (dcomp$media_antag, grupo, mean)
#tapply (dat_media$media_volat, grupo, mean)


dados_final <- bind_cols(cp_data,dageapg) 
dados_final1 <- dados_final %>% arrange(cp1r)


write.csv(dados_final1, "tabela_antagonismo.csv")
```


## Wheather data

```{r}
wheather <- read_excel("wheather-data.xlsx", 2) %>% 
  mutate(Data = as.Date(Data, format= "%Y-%m-%d")) %>% 
  filter(Data >= "2017-10-25" & Data <= "2018-03-15") 

wheather_2017<- read_excel("wheather-data.xlsx", 1) %>% 
  mutate(year = "2017", 
         Data = as.Date(Data, format= "%Y-%m-%d")) %>% 
  filter(Data >= "2017-04-03" & Data <= "2017-09-25")

plot_2018 <- wheather %>% 
  ggplot(aes(x=as.Date(Data))) + 
  geom_line(aes(y = `MinTemp [oC]`*5), color="steelblue", alpha = 0.7) + 
  geom_line(aes(y = `MaxTemp [oC]`*5),  color = "darkred"#, linetype="twodash", linetype='dashed'
            ) +
  #geom_vline(xintercept = as.numeric(as.Date("2017-10-25")), linetype=4) +
  #annotate("text", x = as.numeric(as.Date("2017-10-25")), y = 30, label = "**") +
  geom_bar(aes(y = `Chuva Ac[mm]`), stat = "identity", fill = "grey80") +
  scale_y_continuous(sec.axis = sec_axis(~./5, name = "Temperature (\u00B0C)"),expand=c(0,2))+
  scale_x_date(breaks = "1 month",  labels=date_format("%b/%d"))+
  theme_bw()+
  theme(text = element_text(size = 14), 
       strip.text.x = element_text(size = 14, face = "bold"), 
        )+
  labs(x = "",
       y = "Rainfall (mm)",
       title = "")
  

plot_2017 <- wheather_2017 %>% ggplot(aes(x=as.Date(Data))) + 
  geom_line(aes(y = `MinTemp [oC]`*5), color="steelblue", alpha = 0.7) + 
  geom_line(aes(y = `MaxTemp [oC]`*5),  color = "darkred"#, linetype="twodash"
            ) +
  geom_bar(aes(y = `Chuva Ac[mm]`), stat = "identity", fill = "grey80") +
  scale_y_continuous(sec.axis = sec_axis(~./5, name = "Temperature (\u00B0C)"), expand=c(0,2))+
  scale_x_date(breaks = "1 month",  labels=date_format("%b/%d"))+
  theme_bw()+
  theme(text = element_text(size = 14), 
        strip.text.x = element_text(size = 14, face = "bold"), 
        )+
  labs(x = "",
       y = "Rainfall (mm)",
       title = "") 

combo <- plot_grid(plot_2017, plot_2018, labels=c('2017 season', '2018 season'), align = "hv", ncol=1, rel_heights =  c(1, 1))
ggsave("figs/fig1_wheather.png", combo, width=10, height=4.8, dpi=300)


```
